Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-26 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
updates to plots and fig captions

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}$    & 0.3737                    \\
$\tau^{(0)}$      & 11.2407                   \\
$\omega^{(1)}$    & 0.1258                    \\
$\tau^{(1)}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}$      & $-$1.1010                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
%
\hscomm{I added the values of $p$ to the column headers to keep both columns with the same number of rows. I also
removed the subscript 220 because we already said we consider only this mode.}
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_posteriors_combined.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm cEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
The colors distinguish different $N_{\rm max}$ in the spin expansion.
}
\label{fig:cEFT_exec_sum}
\end{figure}

@@

% We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
% hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
% %
% As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
% %
% Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
% %
% \begin{equation}
%     \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
% \end{equation}
% %
% % \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors' frequency band, making the
parameter estimation challenging.
% \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}~\hscomm{That was the upper prior limit for the GW150914 run. I modified the text to explain this.}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Posterior distribution function for the coupling constant $\ell_{\rm qEFT}$ for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

% \hscomm{Could remove this part until the end of this section.}
%
We can expand the resulting formula in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$ can be
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}, where for the damping time we use the relation ${\rm Im}(\sigma)_{\rm EdGB} = - 1 / \tau_{\rm EdGB}$
and re-expand in $\ell_{\rm EdGB}$ and $\chi$.
%
These steps yield for $p_{\rm EdGB} = 4$,
%
\begin{align}
     \delta \omega_{\rm EdGB}^{(0)} = 0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
\end{align}
%
for the $j=0$ coefficients and
%
\begin{equation}
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{equation}
%
for the $j=1$ coefficients.

@@

We then proceed to determine $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$ as follows.
%
Using the fitting formula Eq.~(54a) of Ref.~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M {\rm Re}(\sigma)_{\rm dCS} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi_{\rm f})^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part, ${\rm Im}(\sigma)_{\rm dCS} =  - 1 / \tau_{\rm dCS}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm dCS}^{4} / (M^4 \kappa)$, and thus
%
\begin{equation}
    \kappa \zeta = (\ell_{\rm dCS} / M)^4 = \gamma_{\rm dCS},
\end{equation}
%
where $c_{i}$ (with $i=1, \dots, 6$) are fitting coefficients which can be found in
Table II of Ref.~\cite{Wagle:2021tam},

>>>>

\begin{table*}[t]
% \begin{tabular}{c | c c c c c c c c c c}
% \hline
% \hline
% Theory      & $\omega^{(0)}_{220}$ & $\tau^{(0)}_{220}$ & $\omega^{(1)}_{220}$ & $\tau^{(1)}_{220}$ & $p$ & $\delta \omega^{(0)}_{220}$ & $\delta \tau^{(0)}_{220}$ & $\delta \omega^{(1)}_{220}$ & $\delta \tau^{(1)}_{220}$ & Ref.  \\
% \hline
% GR          & 0.3737               & 11.2407            &  0.1258              & 0.2522             & --  & --                          & --                        & --                          & --                        & \cite{Maselli:2019mjd} \\
% \hline
% EdGB        & --                   & --                 &  --                  & --                 & 4   & 0.0107                      & 0.0044                    & $-$0.2480                   & $-$1.101                  & \cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} \\
% dCS         & --                   & --                 &  --                  & --                 & 4   & 3.1964                      & 6.3619                    & 41.199                      & 794.66                    & \cite{Wagle:2021tam}  \\
% cubic EFT   & --                   & --                 &  --                  & --                 & 4   & $-$0.5813                   & 2.6469                    & $-$3.8620                   & 265.12                    & \cite{Cano:2021myl} \\
% quartic EFT & --                   & --                 &  --                  & --                 & 6   & $-$0.2114                   & $-$0.6070                 & $-$1.5263                   & 171.35                    & \cite{Cano:2021myl}  \\
% \hline
% \hline
% \end{tabular}
%
\begin{tabular}{c | c }
\hline
\hline
                        & GR~\cite{Maselli:2019mjd} \\
\hline
$\omega^{(0)}$    & 0.3737                    \\
$\tau^{(0)}$      & 11.2407                   \\
$\omega^{(1)}$    & 0.1258                    \\
$\tau^{(1)}$      & 0.2522                    \\
\hline
\hline
\end{tabular}
\quad
\begin{tabular}{c | c  c  c  c}
\hline
\hline
                              & EdGB ($p=4$)~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} & dCS ($p=4$)~\cite{Wagle:2021tam} & cubic EFT ($p=4$)~\cite{Cano:2021myl} & quartic EFT ($p=6$)~\cite{Cano:2021myl} \\
\hline
$\delta\omega^{(0)}$    & 0.0107                                               & 3.1964                   & $-$0.5813                     & $-$0.2114                       \\
$\delta\tau^{(0)}$      & 0.0044                                               & 6.3619                   & $-$0.2114                     & $-$0.6070                       \\
$\delta\omega^{(1)}$    & $-$0.2480                                            & 41.199                   & $-$3.8620                     & $-$1.5263                       \\
$\delta\tau^{(1)}$      & $-$1.1014                                            & 794.66                   & 265.12                        & 171.35                          \\
\hline
\hline
\end{tabular}
\caption{Summary of theory-specific QNM calculations.
%
We summarize each theory we have considered together with: the exponent $p$ at
which their QNM-modification enters, the corresponding modifications to the
oscillation frequency $\delta \omega^{(n)}_{220}$ and decay time $\delta \tau^{(n)}_{220}$, and the
references from which we used the results from.
%
We also included for comparison the GR coefficients, up to $j=1$, obtained in Ref.~\cite{Maselli:2019mjd}.
%
The remaining GR coefficients for $1< j \leqslant 4$, for which their non-GR counterpart cannot be determined as of yet for
the theories under consideration, can be found in Table~I of Ref.~\cite{Maselli:2019mjd}.
%
\hscomm{I added the values of $p$ to the column headers to keep both columns with the same number of rows. I also
removed the subscript 220 because we already said we consider only this mode.}
}
\label{tab:ref_theories_qnms}
\end{table*}

@@

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/ceft_posteriors_combined.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the cubic EFT of GR.
We show our results for
GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
The colors distinguish different $N_{\rm max}$ in the spin expansion and, once more, the vertical
lines mark the 90\% upper credible intervals.
}
\label{fig:cEFT_exec_sum}
\end{figure}

@@

% We find that both the EFT~\eqref{eq:eft_bound} and ParSpec~\eqref{eq:parspec_bound} are satisfied,
% hence allowing us to infer bounds on $\ell_{\rm qEFT}$.
% %
% As Fig.~\ref{fig:qEFT_cdf} shows the CDFs for both events reach 90\% for $\gm = M_{\rm f}$ before $\varepsilon = 1/2$.
% %
% Consequently, from the joint posterior of $\ell_{\rm qEFT}$ we quote our final result,
% %
% \begin{equation}
%     \ell_{\rm qEFT} \leqslant 51.3~\textrm{km} \quad \textrm{at 90\% credible level}\,.
% \end{equation}
% %
% % \abcomm{Please update the paragraph below.}
In this theory, we have considered only $N_{\rm max} = 0$.
%
We find that the addition of spin corrections (while maintaining the same prior
ranges on $\ell_{\rm qEFT}$ as used in the $N_{\rm max} = 0$ study) can result
in waveforms that can have a ringdown segment larger (sometimes seconds long)
than the inspiral-plunge segment in the detectors' frequency band, making the
parameter estimation challenging.
% \abcomm{The reference to 80 km comes out of the blue. Doesn't it? Why did you refer to it?}~\hscomm{That was the upper prior limit for the GW150914 run. I modified the text to explain this.}
% for which the ringdown portions of the larger than the inspiral-plunge part. \abcomm{The last sentence
% has some problems in English. I do not udnerstand it.}
%
To overcome this issue we have lowered the value of $\ell^{\rm max}_{\rm qEFT}$, but by doing so we have obtained posteriors
which were flat, just as our prior, and were thus uninformative. Hence, we do not quote any results for $N_{\rm max} = 1$.
%
Table~\ref{tab:summary_qeft} summarizes our findings for the quartic EFT of GR.
%
\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/qeft_posteriors_combined_corrected.pdf}
\caption{(Color online).~Similar to Fig.~\ref{fig:edgb_exec_sum}, but for the quartic EFT of GR.
We show our results for GW150914 (top panel), GW200129 (middle panel) and combined events (bottom panel).
%
As there, we mark with the vertical lines the 90\% upper credible intervals.
}
\label{fig:qEFT_exec_sum}
\end{figure}

@@

% \hscomm{Could remove this part until the end of this section.}
%
We can expand the resulting formula in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$ can be
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}, where for the damping time we use the relation ${\rm Im}(\sigma)_{\rm EdGB} = - 1 / \tau_{\rm EdGB}$
and re-expand in $\ell_{\rm EdGB}$ and $\chi$.
%
These steps yield for $p_{\rm EdGB} = 4$,
%
\begin{align}
     \delta \omega_{\rm EdGB}^{(0)} = 0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
\end{align}
%
for the $j=0$ coefficients and
%
\begin{equation}
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.1014.
\end{equation}
%
for the $j=1$ coefficients.

@@

We then proceed to determine $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$ as follows.
%
Using the fitting formula Eq.~(54a) of Ref.~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M {\rm Re}(\sigma)_{\rm dCS} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi_{\rm f})^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part, ${\rm Im}(\sigma)_{\rm dCS} =  - 1 / \tau_{\rm dCS}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm dCS}^{4} / (M_{\rm s}^4 \kappa)$, thus
$\kappa \zeta = \gamma_{\rm dCS}$ and
% %
% \begin{equation}
%     \kappa \zeta = (\ell_{\rm dCS} / M^{\rm s})^4 = \gamma_{\rm dCS},
% \end{equation}
% %
where $c_{i}$ (with $i=1, \dots, 6$) are fitting coefficients which can be found in
Table II of Ref.~\cite{Wagle:2021tam},
